---
title: "02. Variance calc."
output:
  html_document:
    df_print: paged
---

This note tries to link the relevant parts for variance calculation between the package implementation and the paper (very rough, cannot guarantee correct)


### Set-up 
```{r}
# Load libraries
library(data.table)
library(mlr3)
library(mlr3learners)
library(DoubleML)

# Load dataset
data = fetch_401k(return_type = "data.table", instrument = TRUE)
# Define covariates
features_base = c("age", "inc", "educ", "fsize", "marr", "twoearn", "db", "pira", "hown")

set.seed(123)

```

\

### DML estimation

```{r}
# Prepare data for DoubleML
data_dml_base = DoubleMLData$new(data,
                                 y_col = "net_tfa",
                                 d_cols = "e401",
                                 x_cols = features_base)

# Initialize LASSO learners
lasso = lrn("regr.cv_glmnet", nfolds = 5, s = "lambda.min")
lasso_class = lrn("classif.cv_glmnet", nfolds = 5, s = "lambda.min")

# Initialize DoubleMLPLR model
dml_plr_lasso_base_n_fold = DoubleMLPLR$new(data_dml_base, 
                                            ml_l = lasso, 
                                            ml_m = lasso_class, 
                                            n_folds = 5,
                                            apply_cross_fitting = TRUE) # to be able to compare without cross fitting (do not accept n_fold >2)

# Fit the model
dml_plr_lasso_base_n_fold$fit()

# Display results
dml_plr_lasso_base_n_fold$summary()
```

\

### Variance calculation

\

See [paper](https://arxiv.org/pdf/1608.00060) for sections that might be relevant

* Section 2.2 (from p13)
    * equ. (2.7) - Neyman orthogonal score
    * equ. (2.8 - 2.10) - solution to the score function


* Section 3.3 (from 28)
    * Theorem 3.3 (p30) - variance (search "and the approximate variance is", see the following line)
       * [psi below  ] psi(W; theta, ita) - corresponds to the psi below (from solving the score function)
       * [psi_a below] J_0 - assumption 3.3 (p28, search "is satisfied, for the Jacobian matrix", see the following line) - should be the first direvative of score function with respect to theta (?)

\

See GitHub implementation at [script](https://github.com/DoubleML/doubleml-for-r/blob/main/R/double_ml.R#L1562)

```{r}
# check the std error (semi-manually)
# ------------------------------------

psi_a <- dml_plr_lasso_base_n_fold$.__enclos_env__$private$get__psi_a()
psi <- dml_plr_lasso_base_n_fold$.__enclos_env__$private$get__psi()
n_obs <- dml_plr_lasso_base_n_fold$data$n_obs


# Apply cross-fitting check
if (dml_plr_lasso_base_n_fold$apply_cross_fitting) {
  scaling_factor <- n_obs
} else {
  # Without cross-fitting, extract test sample indices
  smpls <- dml_plr_lasso_base_n_fold$.__enclos_env__$private$get__smpls()
  test_ids <- smpls$test_ids
  test_index <- test_ids[[1]]  # Use the first fold for simplicity
  psi_a <- psi_a[test_index]
  psi <- psi[test_index]
  scaling_factor <- length(test_index)
}

# Compute the variance and standard error
J <- mean(psi_a)  # Average of psi_a
sigma2_hat <- mean(psi^2) / (J^2) / scaling_factor  # Variance estimate
std_error <- sqrt(sigma2_hat)  # Standard error

cat("Manually calculated standard error:", std_error, "\n")
```


