---
title: "04. Cross-fitting"
output:
  html_document:
    df_print: paged
---
\

This notebook keeps using the Partially linear model (PLR) with linear model as the regression algorithm and logistic model as the classification algorithm and explores different settings for n_fold and cross-fitting application.

With the current data and model specification, sampling can affect coefficient and variance, with bigger difference observed when switching on/ off the cross-fitting.

\

### Setup

```{r}
library(DoubleML)
library(mlr3)
library(mlr3learners)
library(ggplot2)

# Set the logging threshold to "warn" to suppress info messages
lgr::get_logger("mlr3")$set_threshold("warn")

# Prepare the data
data = fetch_401k(return_type = "data.table")
features_base = c("age", "inc", "educ", "fsize", "marr", "twoearn", "db", "pira", "hown")
data_dml_base = DoubleMLData$new(data, y_col = "net_tfa", d_cols = "e401", x_cols = features_base)

# Define linear regression learners for nuisance functions
linear_regression = lrn("regr.lm")       
linear_classification = lrn("classif.log_reg")  
```

\

### Fitting

```{r}

# Define a helper function

dml_n_fold <- function(
  data_input,
  regr_model,
  cls_model,
  n_fold_,
  apply_cross_fitting_,
  output_obj
  
) {
  
  set.seed(123)
  
  dml_learner <- DoubleMLPLR$new(
    data_input,
    ml_l = regr_model,    # Outcome nuisance function
    ml_m = cls_model, # Treatment nuisance function
    n_folds = n_fold_,                 # No folds (use entire data)
    apply_cross_fitting = apply_cross_fitting_  
  )
  
  
  # Fit the model
  dml_learner$fit()
  
  # Update results
  output_obj[['n_fold']] <- append(output_obj[['n_fold']], n_fold_)
  output_obj[['cross_fitting']] <- append(output_obj[['cross_fitting']], as.character(apply_cross_fitting_))
  output_obj[['estimate']] <- append(output_obj[['estimate']], dml_learner$coef)
  output_obj[['std. error']] <- append(output_obj[['std. error']], dml_learner$se)
  output_obj[['lower']] <- append(output_obj[['lower']], dml_learner$confint()[1])
  output_obj[['upper']] <- append(output_obj[['upper']], dml_learner$confint()[2])
  
  return(output_obj)
  
}




# test 11 settings, where
# when n_fold = 1, it cross-fitting is trivial
# when n_fold = 2, we can either use or not use cross-fitting
# when n_fold > 2, can only use cross-fitting

n_folds_setting <- c(1, 2, seq(2, 10))
cross_fit_setting <- c(FALSE, FALSE, rep(TRUE, 9))


# initialise result collector
fit_result <- list(
  'n_fold' = integer(0),
  'cross_fitting' = character(0),
  'estimate' = numeric(0),
  'std. error' = numeric(0),
  'lower' = numeric(0),
  'upper' = numeric(0)
)


# fit
for (index_ in 1:length(n_folds_setting)){
  
  
  fit_result <- dml_n_fold(
    data_input = data_dml_base,
    regr_model = linear_regression,
    cls_model = linear_classification,
    n_fold_ = n_folds_setting[index_],
    apply_cross_fitting_ = cross_fit_setting[index_],
    output_obj = fit_result
  )
  
  
}


# clean up result

fit_result <- data.frame(fit_result)
fit_result
```

\

### Plot

```{r}
# Create the plot
g_ci <- ggplot(fit_result, aes(x = factor(n_fold), y = estimate, color = cross_fitting, group = cross_fitting)) +
  geom_point(position = position_dodge(width = 0.4), size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper), position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed")+
  ggtitle("Coefficients and 0.95 Confidence Interval with PLR") +  # Set the title as the y-axis label 
  labs(
    x = "n_fold",
    color = "Cross Fitting",
    subtitle = "Linear regression model and logistic classifier"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, angle = 0, hjust = 0.5),  # Horizontal tick marks
    axis.title.x = element_text(size = 14),  # Larger x-axis label
    axis.title.y = element_blank(),  # Remove the y-axis label
    legend.position = "bottom",  # Move the legend to the bottom
    text = element_text(size = 14),  # Adjust text sizes for the overall plot
    plot.title = element_text(hjust = 0.5), # put the title in centre
    plot.subtitle = element_text(hjust = 0.5)
  )

# Display the plot
g_ci
```

